---
title: "Data607_Tidyverse_Vignette_Part_2"
author: "C. Rosemond, Extended by Fan Xu"
date: "12/1/2019"
output:
  md_document:
    variant: markdown_github
  html_document:
    df_print: paged
    toc: yes
    toc_collapsed: yes
    toc_float: yes
  pdf_document:
        extra_dependencies: ["geometry", "multicol", "multirow"]
theme: lumen
number_sections: yes
toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Part 1 by Rosemond

title: "Tidyverse Part 1"
author: "C. Rosemond"
date: "November 2, 2019"
output: html_document



## Library

```{r library, message=FALSE}
library(tidyverse)
```


## Data Set(s)

I selected two fivethirtyeight data sets: one that contains current Soccer Power Index (SPI) ratings and rankings for men's club teams and a second that contains match-by-match SPI ratings and forecasts back to 2016.

URL: https://github.com/fivethirtyeight/data/tree/master/soccer-spi


## readr - read_csv()

The readr package facilitates the reading in of 'rectangular' data like .csv files or other delimited files. Here, I use the read_csv() function to read in two data sets: the global rankings, or 'rankings', and the matches, or 'matches'.

```{r read_csv1, eval=TRUE, message=FALSE}
rankings <- read_csv('https://projects.fivethirtyeight.com/soccer-api/club/spi_global_rankings.csv')
matches <- read_csv('https://projects.fivethirtyeight.com/soccer-api/club/spi_matches.csv')
head(rankings)
tail(matches)
```


## dplyr - mutate()

The dplyr package provides a grammar for the manipulation of data--notably, in data frames or tibbles. Here, I use the mutate function to add a new column--a match ID--to the matches tibble.


```{r mutate, eval=TRUE}
matches <- mutate(matches, match = row_number())
matches <- matches[,c(23,1:22)]
head(matches)
```


## dplyr - select()

The select function from dplyr enables the selection of data frame columns by name or helper function. Here, I select and keep the first six columns ('match' through 'team2') from the matches tibble.

```{r select, eval=TRUE}
matches <- select(matches, match:team2)
head(matches)
```


## dplyr - filter()

The filter function from dplyr enables the subsetting of rows based on specified logical criteria. Here, I select matches that occurred from November 1st through November 7th.

```{r filter, eval=TRUE}
matches <- filter(matches, date >= '2019-11-01' & date <= '2019-11-07')
head(matches)
```


## tidyr - gather()

The tidyr package is designed to facilitate reshaping data. Here, I use the gather() function to reshape the matches tibble from wide to long format, gathering the separate team columns.

```{r gather, eval=TRUE}
matches <- matches %>% gather(-match, -date, -league_id, -league, key=team_number, value=name) %>% select(-team_number)
head(matches)
```


## dplyr - arrange()

The arrange function from dplyr enables the sorting of data based upon column values. Here, I arrange the matches tibble by match number.

```{r arrange, eval=TRUE}
matches <- arrange(matches, match)
head(matches)
```


## dplyr - left_join()

The left_join function works similarly to its SQL counterparts. I finish by using 'name' to merge the matches tibble with the rankings tibble, which contains club rankings and ratings as of November 7th.

```{r left_join, eval=TRUE}
merged <- dplyr::left_join(matches, rankings, by='name')
merged <- select(merged, -league.y, - off, -def)
head(merged)
```


# Part 2 Extension by Fan Xu

title: "Data607_Tidyverse_Vignette_Part_2"
author: "Fan Xu"
date: "12/1/2019"


## stringr::str_extract

`str_extract` is used when I want to get content from a string according to a comment pattern denoted by regular expression. Regular expression is out of the scope of this assignment so I won't go further into details. In this example I used `str_extract` the first 4 digit combo in column 'date' which is the year.

```{r str_extract}
date_year <- mutate(merged, year = str_extract(date, '[0-9]{4}'))
date_year
```


## tidyr::seperate

Another way to get the year column is to convert the column into 'Date' format then use `seperate` function to seperate 'year', 'month' and 'date' into three columns.

```{r separate}
date_year <- mutate(merged, date = as.Date(date))
date_year <- separate(date_year, date, c('year','month','date'))

date_year
```


## tidyr::unite

The `unite` function is used to comcatnate values of two columns into a new column. The argument 'remove' is to indicate whether to remove the original columns that to be united. 

```{r unite}
date_unite <- unite(date_year, year_month, year, month, sep = '-', remove = TRUE)
date_unite
```


## dplyr::group_by & dplyr::tally

The `group_by` function groups the dataframe into groups to allow the following operations to be performed by group. Here, I use `group_by` to group the data by the column `league` and `team_name`, then use `tally` to take a look at how many times each team shows up.

```{r group_by, eval=TRUE, message=FALSE}
group <- group_by(date_unite, league.x, name)
group <- tally(group)
group <- arrange(group, n)
group
```


## Pipe Operator %>%  &  dplyr::rename

The pipe operator '%>%',comes from the `magrittr` package, are embeded in `tidyverse`. It enables a handy coding by inserting the output of the preceding code into the following code by the operator '%>%', which makes R coding cleaner, easier and more straightforward visually. I can perform similar actions from the above code chunks into one piece of code as below:

Note that I used the original spi_matches.csv to extract my desired dataset.

```{r pipe, eval=TRUE, message=FALSE}

# To generate a new dataset from spi_matches.csv
matches_orig <- read_csv('https://projects.fivethirtyeight.com/soccer-api/club/spi_matches.csv')

overall <- matches_orig %>%
  mutate(match = row_number()) %>%
  .[,c(23,1:22)] %>%
  select(match:spi2) %>%
  filter(team1=='Liverpool' | team1=='Arsenal' | team1=='Barcelona') %>%
  select(-team2, -spi2) %>%
  arrange(date) %>%
  rename(name = team1, SPI = spi1) %>%
  left_join(rankings, by='name') %>%
  select(-league.y, - off, -def, -spi, -rank, -prev_rank) %>%
  separate(date, c('year','month','date')) %>%
  unite(month_date, month, date, sep = '-', remove = TRUE) %>%
  mutate(year = as.numeric(year))
  
  
overall
```


## Visualization: ggplot2

`ggplot2` is a classific package for ploting the data. 

In this example I plot a line graph with column `full_date` as x-axis and `SPI` as y-axis.  The graph below shows the averaged forecasted SPI values from 2016 to 2020 for the teams Liverpool, Arsenal and Barcelona.

```{r plot}
overall %>% 
  group_by(name, year) %>%
  summarise(avg_SPI = mean(SPI)) %>%
  ggplot() +
  geom_point(aes(x = year, y = avg_SPI)) +
  geom_line(aes(x = year, y = avg_SPI, colour = name)) +
  ggtitle('SPI Forecast from 2016 to 2020') +
  xlab('Year') +
  ylab('Average SPI')
```

