---
title: "Tidyverse Assignment Part 02 Extended from SHaslett-607-Tidyverse-assignment.Rmd"
author: "Forhad Akbar"
date: "11/27/2019"
output:
  rmdformats::readthedown
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment Overview

**Create an Example** Using one or more TidyVerse packages, and any dataset from fivethirtyeight.com or Kaggle, create a programming sample “vignette” that demonstrates how to use one or more of the capabilities of the selected TidyVerse package with your selected dataset.

```{r, eval=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(rmdformats)
```


## Assignment Response

For this assignment, I have chosen to demonstrate the basic functionailty supplied by the dplry package that comes with tidyverse.

## Datasource
https://www.kaggle.com/residentmario/ramen-ratings/data


#### Import the Ramen Ratings dataset using the "readr" package.


```{r, eval=TRUE, message=FALSE, warning=FALSE}
ramen_ratings <- read.csv("https://raw.githubusercontent.com/stephen-haslett/FALL2019TIDYVERSE/master/ramen-ratings.csv", header = TRUE)
```


#### Check that the data imported successfully.

```{r, eval=TRUE, message=FALSE, warning=FALSE}
ramen_ratings <- as_tibble(ramen_ratings)
head(ramen_ratings)
```


#### Using the "dplyr" package's "slice" function, select a specific range of rows from the dataset.

```{r, eval=TRUE, message=FALSE, warning=FALSE}
# Select rows 10 to 20 from the dataset.
selected_rows <- slice(ramen_ratings, 10:19)
selected_rows
```


#### Now use the "slice" function to select 2 row ranges, and 4 specific rows.

```{r, eval=TRUE, message=FALSE, warning=FALSE}
# Select rows 2 to 9, 26 to 30, and rows 40, 21, 16, and 35.
multiple_selected_rows <- ramen_ratings %>% slice(c(2:9, 26:30, 40, 21, 16, 35))
multiple_selected_rows
```


#### Use dplyr's "filter" function to select rows based on specified conditions.

_In this example, we will only select rows that have a 5 Star rating._

```{r, eval=TRUE, message=FALSE, warning=FALSE}
five_stars <- filter(ramen_ratings, Stars == 5)
five_stars
```


#### Finally, use dplyr's "select" function to select specific data columns from the dataset.

_In this example, we will select the "Brand", "Variety", "Country", "Stars", and "Top.Ten" columns for brands that have a Top.Ten listing._

```{r, eval=TRUE, message=FALSE, warning=FALSE}
# Select only the rows with Top.Ten column entries whilst removing those
# with NA, empty, or "\n" values.
ramen_ratings_filtered <- filter(ramen_ratings, Top.Ten != "" & !is.na(Top.Ten) & Top.Ten != "\n")

ramen_top_ten <- select(ramen_ratings_filtered, c("Brand", "Variety", "Country", "Stars", "Top.Ten"))
ramen_top_ten
```
#### ===========================================================================================

#### Extended SHaslett-607-Tidyverse-assignment.Rmd By Forhad Akbar

#### Added dplyr function group_by().

`group_by` capability tutorial with tally

Description: Using group_by and tally we can find count of category member  
Usage: group_by(data, …)  
Example: Find the count by Ramen Style

```{r}
style <- ramen_ratings %>% group_by(Style) %>% tally()
style
```

#### Added tidyverse function ggplot().

```{r}
ggplot(style, aes(x = reorder(Style,n), y = n)) +
    geom_bar(stat = "identity", position = position_dodge(),  fill="steelblue") +
    geom_text(aes(label = n), vjust = .5, hjust = 1, position = position_dodge(width = 0.9), color = "black") +
        ggtitle("Ramen Style By Count") +
    xlab("Style") + ylab("Count") +
        coord_flip()
```