---
title: "Tidyverse Assignment - Ramen Ratings"
author: "Stephen Haslett"
date: "11/25/2019"
output:
  rmdformats::readthedown
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment Overview

**Create an Example** Using one or more TidyVerse packages, and any dataset from fivethirtyeight.com or Kaggle, create a programming sample “vignette” that demonstrates how to use one or more of the capabilities of the selected TidyVerse package with your selected dataset.

```{r, eval=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(rmdformats)
```


## Assignment Response

For this assignment, I have chosen to demonstrate the basic functionailty supplied by the dplry package that comes with tidyverse.

## Datasource
https://www.kaggle.com/residentmario/ramen-ratings/data


#### Import the Ramen Ratings dataset using the "readr" package.


```{r, eval=TRUE, message=FALSE, warning=FALSE}
ramen_ratings <- read.csv("https://raw.githubusercontent.com/stephen-haslett/FALL2019TIDYVERSE/master/ramen-ratings.csv", header = TRUE)
```


#### Check that the data imported successfully.

```{r, eval=TRUE, message=FALSE, warning=FALSE}
ramen_ratings <- as_tibble(ramen_ratings)
head(ramen_ratings)
```


#### Using the "dplyr" package's "slice" function, select a specific range of rows from the dataset.

```{r, eval=TRUE, message=FALSE, warning=FALSE}
# Select rows 10 to 20 from the dataset.
selected_rows <- slice(ramen_ratings, 10:19)
selected_rows
```


#### Now use the "slice" function to select 2 row ranges, and 4 specific rows.

```{r, eval=TRUE, message=FALSE, warning=FALSE}
# Select rows 2 to 9, 26 to 30, and rows 40, 21, 16, and 35.
multiple_selected_rows <- ramen_ratings %>% slice(c(2:9, 26:30, 40, 21, 16, 35))
multiple_selected_rows
```


#### Use dplyr's "filter" function to select rows based on specified conditions.

_In this example, we will only select rows that have a 5 Star rating._

```{r, eval=TRUE, message=FALSE, warning=FALSE}
five_stars <- filter(ramen_ratings, Stars == 5)
five_stars
```


#### Finally, use dplyr's "select" function to select specific data columns from the dataset.

_In this example, we will select the "Brand", "Variety", "Country", "Stars", and "Top.Ten" columns for brands that have a Top.Ten listing._

```{r, eval=TRUE, message=FALSE, warning=FALSE}
# Select only the rows with Top.Ten column entries whilst removing those
# with NA, empty, or "\n" values.
ramen_ratings_filtered <- filter(ramen_ratings, Top.Ten != "" & !is.na(Top.Ten) & Top.Ten != "\n")
ramen_top_ten <- select(ramen_ratings_filtered, c("Brand", "Variety", "Country", "Stars", "Top.Ten"))
ramen_top_ten
```

# This is what Bryan Persaud is adding to Stephen Haslett's code

```{r}
ggplot(ramen_top_ten, aes(x = Brand, y = Stars)) + geom_bar(stat = "identity") + labs(title = "Top Ten Listings by Stars", x = "Brand", y = "Stars") + theme(axis.text.x = element_text(angle = 75, size = 14, hjust = 1), axis.text.y = element_text(size = 10))
```

```{r}
ggplot(five_stars, aes(x = Brand, y = Country)) + geom_bar(stat = "identity") + labs(title = "Five Star Ramens by Country", x = "Brand", y = "Country") + theme(axis.text.x = element_text(angle = 75, size = 14, hjust = 1), axis.text.y = element_text(size = 10))
```

I chose to use ggplot2 package to plot a graph to show the ramens that got a top ten listings from the ramen_top_ten dataframe created. I also used ggplot2 to show the ramens that got a five star rating from the five_stars dataframe.