---
title: "TidyVerse Assignment Part 01"
author: "Forhad Akbar"
date: "11/27/2019"
output:
  prettydoc::html_pretty:
    pdf_document: default
    theme: cayman
  pdf_document:
    number_sections: no
  html_document:
    df_print: paged
---

The problem statement is as follows:

In this assignment, you’ll practice collaborating around a code project with GitHub.  You could consider our collective work as building out a book of examples on how to use TidyVerse functions.

GitHub repository:  https://github.com/acatlin/FALL2019TIDYVERSE 

FiveThirtyEight.com datasets.

Kaggle datasets. 

You have two tasks:

Create an Example.  Using one or more TidyVerse packages, and any dataset from fivethirtyeight.com or Kaggle, create a programming sample “vignette” that demonstrates how to use one or more of the capabilities of the selected TidyVerse package with your selected dataset. (25 points)
Extend an Existing Example.  Using one of your classmate’s examples (as created above), extend his or her example with additional annotated code. (15 points)
You should clone the provided repository.  Once you have code to submit, you should make a pull request on the shared repository.  Minimally, you should be submitted .Rmd files; ideally, you should also submit an .md file and update the README.md file with your example.

After you’ve completed both parts of the assignment, please submit your GitHub handle name in the submission link provided in the week 1 folder!  This will let your instructor know that your work is ready to be graded.

You should complete both parts of the assignment and make your submission no later than the end of day on Sunday, December 1st.

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library('knitr')
```

## Data Source

https://www.kaggle.com/ronitf/heart-disease-uci

## Load Data

```{r Load Data}
heart <- read_csv("https://raw.githubusercontent.com/forhadakbar/data607fall2019/master/Week%2014/heart.csv")
kable (head(heart))
```

## Capability 1.

`slice` capability tutorial

Description: Using slice we can select rows by specifying the row number.  
Usage: slice(data, …)  
Example: To select rows 6 to 12


```{r}
slice(heart, 6:12)
```

OR

To select rows 10 to 15, 18 and 299 to 302

```{r}
heart %>% slice(c(10:15, 18, 299:302))
```

## Capability 2.

`mutate` capability tutorial

Description: Creates new columns based on existing ones  
Usage: mutate(.data, …)  
Example: Let’s look at a ratio of resting blood pressure to cholesterol under a new column name “Ratio”

```{r}
#Ratio = trestbps/chol
heart %>% select(trestbps, chol) %>% mutate(Ratio = trestbps/chol)
```

## Capability 3.

`summarise` capability tutorial

Description: Applies functions that return results of length 1. Can perform multiple calculations in the same call.  
Usage: summarise(data, …)  
Example: Lets see mean and median cholesterol along with mean and median maximum heart rate achieved.

```{r}
heart %>% summarise(Avg_cholesteral = mean(chol), Median_cholesteral = median(chol), Avg_heartrate = mean(thalach), Median_heartrate = median(thalach))
```

## Capability 4.

`group_by` capability tutorial with tally

Description: Using group_by and tally we can find count of category member  
Usage: group_by(data, …)  
Example: Find the count by sex

```{r}
heart %>% group_by(sex) %>% tally()
```
