---
title: "Tidyverse Assignment Part 02 extended from purrr - Euclid Zhang.RMD"
author: "Shovan Biswas"
date: "11/27/2019"
output:
  pdf_document: default
  html_document: default
---

The purrr package is a convenient tool for manipulating lists with functions.
This assignment is a demonstration of how to use the functions in the purrr package.

Load the purrr library and the dplyr library as an assistant. We will use the college_all_ages data set from the fivethirtyeight package.
```{r}
library(knitr)
library(tidyverse)
library(fivethirtyeight)
```

First, let look at the first few rows of the college_all_ages data set
```{r}
kable(head(college_all_ages))
```

Extract the unemployment_rate column as a list
```{r}
major_unemp <- college_all_ages$unemployment_rate
```

The set_names function is used to assign a name to each of the element in a list.  
In this example, I assign the names by their corresponding majors
```{r}
major_unemp <- set_names(major_unemp,college_all_ages$major)
head(major_unemp)
```
  
  
  
  

Section 1 - select particular elements from a list:  
1.pluck() ---- select an element by name or index.  
For example: We can select the first element by
```{r}
pluck(major_unemp, "General Agriculture")
# or
pluck(major_unemp, 1)
```

2.keep() ---- select all elements that pass a logicl test.  
For example: We can select all elements whose value is greater than 0.05
```{r}
res <- keep(major_unemp, function(x) x > 0.05)
head(res)
```

3.discard() ---- select all elements that DO NOT pass a logicl test.  
For example: We can select all elements whose value is NOT greater than 0.05
```{r}
res <- discard(major_unemp, function(x) x > 0.05)
head(res)
```

4.head_while() ---- select all elements from the beginning of a list up to and exclude the element that does not pass a logical test.  
For example: we can select all elements whose value is greater than 0.02 from the beginning of a list
```{r}
res <- head_while(major_unemp, function(x) x > 0.02)
head(res)
```

5.tail_while() ---- select all elements from the end of a list up to and exclude the element that does not pass a logical test.  
For example: we can select all elements whose value is greater than 0.05 from the end of a list
```{r}
res <- tail_while(major_unemp, function(x) x > 0.05)
head(res)
```
  
    
  
  
  
Section 2 - logical operations:  
1.every() ---- Check if all elements in a list pass a logical test  
For example: we can check if all elements are greater than 0
```{r}
every(major_unemp,function(x) x > 0)
```

2.some() ---- Check if some elements in a list pass a logical test  
For example: we can check if some elements are 0
```{r}
some(major_unemp,function(x) x == 0)
```

3.some() ---- Check if a value is an element of a list  
For example: we can check if 0 is in the list or not
```{r}
has_element(major_unemp,0)
```

4.detect() ---- Find the first element that passes a logical test  
For example: we can find the first element that is greater than 0.08
```{r}
detect(major_unemp,function(x) x > 0.08)
```

5.detect_index() ---- Find the index of the first element that passes a logical test  
For example: we can find index of the first element that is greater than 0.08
```{r}
detect_index(major_unemp,function(x) x > 0.08)
```
  
    
  
  
  
Section 3 - applying function conditionally:  
1.modify() ---- Apply a function to each of the elements in a list  
For example: multiply all values in the list by 100
```{r}
res <- modify(major_unemp, function(x) x * 100)
head(res)
```

2.modify_at() ---- Apply a function to an element by name or index  
For exmaple: multiply the first element by 100
```{r}
res <- modify_at(major_unemp,1, function(x) x * 100)
head(res)
```

3.modify_if() ---- Apply a function to elements that pass a logical test    
For exmaple: multiply the elements that are greater than 0.05 by 100
```{r}
res <- modify_if(major_unemp, function(x) x > 0.05, function(x) x *100)
tail(res)
```
  
    
  
  
  
Section 4 - lists combine:  
1.append() ---- Append another list to the end of a list  
For example: Append the 3rd and 4th elements to the 1st and 2nd elements as a list
```{r}
append(major_unemp[1:2], major_unemp[3:4])

```

2.prepend() ---- Add another list to the beginning of a list    
For example: Add the 3rd and 4th elements to the beginning of the 1st and 2nd elements as a list
```{r}
prepend(major_unemp[1:2], major_unemp[3:4])
```

3.splice(): combine all lists and other elements into a list  
For example: Combine the 1st and 2nd elements as a list, the 3rd and 4th elements as a list and a new element "0.123" into one list
```{r}
splice(major_unemp[1:2], major_unemp[3:4], "new value" = "0.123")
```

    
  
  
  
Other functions:
  
cross2() ---- Using elements from 2 lists, create a list of all possible combinations with 1 element for the first list and another element from the second list  
For example: create all combinations by using the first 2 values from the major column and the first 2 values from the unemployment rate column
```{r}
cross2(college_all_ages$major[1:2],college_all_ages$unemployment_rate[1:2])
```

array_tree() ---- Turn an array into a list
```{r}
class(major_unemp)
res <- array_tree(major_unemp)
class(res)
```


map2() ---- apply a function to each pair of elements from 2 lists and return the results as a list  
For example: we combine 1 element from the major_category list and a corresponding element from the unemployment_rate list into a list. The result will be a list of 2-element lists.
```{r}
major_unemp_cate <- map2(college_all_ages$major_category, college_all_ages$unemployment_rate, function(x,y) list(x,y))
major_unemp_cate <- set_names(major_unemp_cate,college_all_ages$major)
head(major_unemp_cate)
```

invoke_map() ---- given a list of functions and a list of elements of the same length, apply each of the function to a corresponding element. 
For example: we can multiply the first element by 100 and multiply the second element by -1
```{r}
res <- invoke_map(list(function(x) x * 100, function(y) y * -1), major_unemp[1:2])
head(res)
```


flatten() ---- remove a level of indexes from a list  
For example, we can flatten the just created list of 2-element lists into a single list
```{r}
length(major_unemp_cate)
res <-flatten(major_unemp_cate)
length(res)
```

transpose() ---- Transposes the index order in a multi-level list  
For example, we can Transposes just created the list of 2-element lists 
```{r}
length(major_unemp_cate)
res <- transpose(major_unemp_cate)
length(res)
```

reduce() ---- apply a functin recursively to each element of a list  
For example: we can add up all elements in a list recursively
```{r}
reduce(major_unemp, sum)
```

accumulate() ---- the same as reduce(), but also return all intermediate results as a list  
```{r}
res <- accumulate(major_unemp, sum)
tail(res)
```

#### ===========================================================================================

#### Extended purrr - Euclid Zhang.RMD By Shovan Biswas

#### Added dplyr function select().

`select` capability tutorial

Description: Using select we can keep the selected variables  
Usage: select(data, …)  
Example: To keep only major_category, total we can pass college_all_ages to the function.


```{r}
major_category1 <- college_all_ages %>% select(major_category, total)
major_category1
```


#### Added dplyr function summarise().


`summarise` capability tutorial

Description: Summarizes numerics, over another parameter.  
Usage: summarise(data, …)  
Example: Summarizing total over major_category.



```{r}
total1 <- major_category1 %>% group_by(major_category) %>% summarise(Total = sum(total))
kable(total1)
```

#### Added tidyverse function top_n().

`top_n` capability tutorial

Description: top_n selcts top_n(n) number of records.  
Usage: top_n(n)  
Example: top_n(10.


```{r}
topten <- college_all_ages %>% group_by(major) %>% summarise(Employed = sum(employed)) %>% top_n(10)
topten
```

#### Added tidyverse function ggplot().


`ggplot` capability tutorial



```{r}
ggplot(topten, aes(x = reorder(major,Employed), y = Employed)) +
    geom_bar(stat = "identity", position = position_dodge(),  fill="steelblue") +
    geom_text(aes(label = Employed), vjust = .5, hjust = 1, position = position_dodge(width = 0.9), color = "black") +
        ggtitle("Number of students, employed by top_(10) major") +
    xlab("major") + ylab("employed") +
        coord_flip()
```
