---
title: "Tidyverse example"
author: "Scott Reed"
date: "11/30/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(purrr)
```

## Tidyverse Example -- readr - dplyr - purrr - stringr  - Fight Songs

This is a short example to show some tidyverse functionality. We first use readr to grab the FiveThirtyEight fight song database of categorized fight songs, and force it to be a tibble. 

```{r }
fightSongs<-read_csv(url("https://raw.githubusercontent.com/fivethirtyeight/data/master/fight-songs/fight-songs.csv"))
```

## Dplyr to clean data

We have some unknowns and those will best be handled by marking those as NA. 

```{r}
fightSongs<-fightSongs %>% mutate_if(is.character, list(~na_if(., "Unknown")))
fightSongs$year <- as.numeric(fightSongs$year)
```

We can then use ggplot and a simple dplyr filter to see a plot of when songs were written
```{r}
fightSongs %>% filter(!is.na(year)) %>% ggplot()+geom_bar(aes(year - year %% 10, group=conference, color=conference, fill=conference))
```

We can also see if names that are longer (using stringr to split on white space, and the "friendly with tidyverse) are more likely to have songs that spell out names. 

```{r}
fightSongs$nameLength <-unlist( purrr::map(purrr::map(str_split(fightSongs$writers,"\\s"), nchar),max))
fightSongs  %>% filter(!is.na(nameLength)) %>% ggplot(aes(x=spelling, fill=nameLength>7)) + geom_bar(position="dodge", stat="count")
```

Well it does seem that there may be more spelling with longer names.

## Summarizing by Duration
First, we'll take a look at the the average number of seconds per song, grouped by student_writer:

``` {r}

library(tidyverse)

fightSongs %>%
  group_by(student_writer) %>%
  summarise(avg_seconds = mean(sec_duration),
            median_seconds = median(sec_duration),
            sd_seconds = sd(sec_duration))

boxplot(sec_duration~ student_writer,data = fightSongs)

```

It's a bit hard to see from the boxplot, but we can tell from the summary data that student writers tend to write songs that are around 5 seconds longer than those not written by student writers. 

Next, we'll look at the bpm, also grouped by student_writer:
``` {r}

fightSongs %>%
  group_by(student_writer) %>%
  summarise(avg_bpm = mean(bpm),
            median_bpm = median(bpm),
            sd_bpm = sd(bpm))

boxplot(bpm~ student_writer,data = fightSongs)

```

Interestingly, student writers tend to write songs that are slower than non-student writers!