---
title: "How to use forcat-tidyverse"
author: "Salma Elshahawy"
date: "11/15/2019"
output: html_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Forcats package

forcats provides a suite of useful tools that solve common problems with factors. R uses factors to handle categorical variables, variables that have a fixed and known set of possible values.

In the following dataset, we have some categorical variables like **type_of_subject**, **subject_race**. R uses factors to handle those kinds of variables that have  a fixed set of possible values. The forcats package goal is to provide a convenient tools that can solve some issues when dealing with factors, for example, changing the order of  levels or the  values. The following vignette will demonestrate more.

```{r message=FALSE, warning=FALSE}
if (!require('readr')) install.packages ('readr')
if (!require('forcats')) install.packages ('forcats')
if (!require('dplyr')) install.packages ('dplyr')
if (!require('kableExtra')) install.packages ('kableExtra')
if (!require('ggplot2')) install.packages ('ggplot2')
```

The following dataset is a collection of imdb movies including some data about release dat, country of origin,..etc. The dataset was imported from the fivethirtyeight github repo.

I used another useful tidyverse package called **readr** which gives the flexibility to import different file format into your r workspace.

```{r}
url <- "https://raw.githubusercontent.com/fivethirtyeight/data/master/biopics/biopics.csv"
df <- read.csv(url, header = TRUE)
head(df)
```

If I tried to plot the **type_of_subject** variable as the illustrated below, you will find that the bars haven't distributed uniformly even if you forced them to reorder.  

```{r}
df_2 <- df %>%
  arrange(type_of_subject)

ggplot(df_2, aes(x = type_of_subject)) + 
  geom_bar() + 
  coord_flip()
```

To solve this problem, you can use the **fct_infreq()** wich is ships within the **forcats** package. This function is responsible for reordering a factor variable by the frequency of value - notice the change below:

```{r}
df %>%
  mutate(type_of_subject = fct_infreq(type_of_subject)) %>%
  ggplot(aes(x = type_of_subject)) + 
  geom_bar() + 
  coord_flip()
```

Pretty neat! you can easily identify the trends from graphs without any extra processing. 














