---
title: "Tidyverse - Zhi Ying Chen"
output: 
  html_document:
    df_print: paged
  md_document:
    variant: markdown_github
---
#Part I
#Using one or more TidyVerse packages, and any dataset from fivethirtyeight.com or Kaggle, create a programming sample "vignette" that demonstrates how to use one or more of the capabilities of the selected TidyVerse package with your selected dataset. (25 points)
```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
```
#Load the data from Kaggle
```{r}
avocado_price <- read.csv("https://raw.githubusercontent.com/Zchen116/tidyverse/master/avocado.csv")
head(avocado_price)
summary(avocado_price)
```
#Use filter function to filter Boston's avocado price from the dataset
```{r}
head(filter(avocado_price,region== "Boston"))
```
#Use arrange function to arrange the Average Price from the highest
```{r}
head(avocado_price%>%arrange(desc(AveragePrice)))
```
#Use arrange function to arrange the Average Price from the lowest
```{r}
head(avocado_price%>%arrange(AveragePrice))
```
#Use select function to choose data of region, year, type, and Average Price to create another data frame
```{r}
head(avocado_price%>%select(region, year,type,AveragePrice))
```
#Use rename function to rename column
```{r}
head(rename(avocado_price, city = region))
```
#Use summarise function to a data frame  about minium, mean, and maxium of Average Price
```{r}
filter(avocado_price,region == "Boston")%>%summarise(AveragePrice_min=min(AveragePrice), AveragePrice_mean=mean(AveragePrice),AveragePrice_max=max(AveragePrice))
```
#Use group_by function
```{r}
by_region <- group_by(avocado_price, region)
sui <- summarise(by_region,
  count = n(),
  AveragePrice_mean = mean(AveragePrice, na.rm = TRUE))
head(sui %>% arrange(desc(AveragePrice_mean)))
```
#Use mutate function
```{r}
my_data<-head(sui %>% arrange(desc(AveragePrice_mean)) %>%  mutate(AveragePrice_mean = round(AveragePrice_mean, 2)),20)
my_data
```
#Use ggplot2 function to create graph
```{r}
my_data%>%ggplot(aes(x=region, y=AveragePrice_mean, fill=region))+
  geom_bar(stat = "identity", position = "dodge") + 
  guides(fill = FALSE) +
  ggtitle("Average Price mean") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

### Lin Li's extension to Part 1
```{r}
by_year <- group_by(avocado_price, year)
sui2 <- summarise(by_year, AvgPrice_mean = mean(AveragePrice, na.rm = TRUE))
head(sui2 %>% arrange(year))

```

```{r}
PriceByYear <- ggplot(data = sui2, aes(x=year, y=AvgPrice_mean, fill=year))+
  geom_bar(stat = "identity", position = "dodge") +
  ggtitle("Mean Price by Year") 
  
PriceByYear

```


#Part 2
#Extend an Existing Example.  Using one of your classmate's examples (as created above), extend his or her example with additional annotated code. (15 points)
#1, Get dataset from 538 (existing code)
```{r}
weather <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/us-weather-history/KCLT.csv")
summary(weather)
```
#2, Tidyr separate function (existing code):
```{r}
library(tidyr)
weather2 <- weather %>% separate(date, c("year", "month", "day"), sep = "-")
summary(weather2)
```
#3, Dplyr select function (existing code):
```{r}
head(select(weather2, year, actual_mean_temp, record_min_temp, record_max_temp, record_precipitation))
```
#4, Dplyr filter (subsetting dataset) (existing code):
```{r}
head(filter(weather2, year == "2014"))
```
#Extend Part:
#Use select and filter funtion 
```{r}
weather3 <- weather2 %>%select(year, actual_mean_temp, actual_min_temp, actual_max_temp, actual_precipitation)
head(filter(weather3, year == "2014"))
```
#Use summarise function to a data frame  about minium, mean, and maxium
```{r}
filter(weather3, year == "2014")%>%summarise(min=min(actual_min_temp), mean=mean(actual_mean_temp),max=max(actual_max_temp))
```



