---
title: "Tidyverse Part 2"
author: "Salma Elshahwy"
date: "December 1, 2019"
output: html_document
---
### This is extension to C. Rosemond example

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Library

```{r library, message=FALSE}
library(tidyverse)
```

## Data Set(s)
I selected two fivethirtyeight data sets: one that contains current Soccer Power Index (SPI) ratings and rankings for men's club teams and a second that contains match-by-match SPI ratings and forecasts back to 2016.

URL: https://github.com/fivethirtyeight/data/tree/master/soccer-spi

## readr - read_csv()
The readr package facilitates the reading in of 'rectangular' data like .csv files or other delimited files. Here, I use the read_csv() function to read in two data sets: the global rankings, or 'rankings', and the matches, or 'matches'.

```{r read_csv1, eval=TRUE, message=FALSE}
rankings <- read_csv('https://projects.fivethirtyeight.com/soccer-api/club/spi_global_rankings.csv')
matches <- read_csv('https://projects.fivethirtyeight.com/soccer-api/club/spi_matches.csv')
head(rankings)
tail(matches)
```


## dplyr - mutate()
The dplyr package provides a grammar for the manipulation of data--notably, in data frames or tibbles. Here, I use the mutate function to add a new column--a match ID--to the matches tibble.

```{r mutate, eval=TRUE}
matches <- mutate(matches, match = row_number())
matches <- matches[,c(23,1:22)]
head(matches)
```


## dplyr - select()
The select function from dplyr enables the selection of data frame columns by name or helper function. Here, I select and keep the first six columns ('match' through 'team2') from the matches tibble.

```{r select, eval=TRUE}
matches <- select(matches, match:team2)
head(matches)
```


## dplyr - filter()
The filter function from dplyr enables the subsetting of rows based on specified logical criteria. Here, I select matches that occurred from November 1st through November 7th.

```{r filter, eval=TRUE}
matches <- filter(matches, date >= '2019-11-01' & date <= '2019-11-07')
head(matches)
```


## tidyr - gather()
The tidyr package is designed to facilitate reshaping data. Here, I use the gather() function to reshape the matches tibble from wide to long format, gathering the separate team columns.

```{r gather, eval=TRUE}
matches <- matches %>% gather(-match, -date, -league_id, -league, key=team_number, value=name) %>% select(-team_number)
head(matches)
```


## dplyr - arrange()
The arrange function from dplyr enables the sorting of data based upon column values. Here, I arrange the matches tibble by match number.

```{r arrange, eval=TRUE}
matches <- arrange(matches, match)
head(matches)
```


## dplyr - left_join()
The left_join function works similarly to its SQL counterparts. I finish by using 'name' to merge the matches tibble with the rankings tibble, which contains club rankings and ratings as of November 7th.

```{r left_join, eval=TRUE}
merged <- dplyr::left_join(matches, rankings, by='name')
merged <- select(merged, -league.y, - off, -def)
head(merged)
```

## tidyverse - ggplot()
ggplot2 is a system for declaratively creating graphics, based on The Grammar of Graphics. You provide the data, tell ggplot2 
how to map variables to aesthetics, what graphical primitives to use, and it takes care of the details.
in this example, I wanted to see the distribution of ranks vs. prev_ranks in different dates.

```{r}
# we can use ggplot to have a better sense of the analysis
library(ggplot2)
ggplot(merged, aes(rank, prev_rank)) + 
  geom_point() +
  facet_wrap(~date)
```





