---
title: "Tidyverse_assignment_Task1"
author: "Jayanth0572"
date: "November 25, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview
In this task dataset "breweries_us.csv" was downloaded from "https://www.kaggle.com/brkurzawa/us-breweries/data#".</br>
This CSV file was imported into R and few transacformations were performed using packages from Tidyverse.</br>
Below packages from Tidyverse was used...

- readr
- stringr
- magrittr


# Code
### Read dataset into R
using readr package
```{r}

library(readr)

# Source file "breweries_us.csv" was downloaded from "https://www.kaggle.com/brkurzawa/us-breweries/data#"

# Read data from csv file "breweries_us.csv" into a tibble
# "read_csv" function from "readr" package was sed
input <- read.csv("breweries_us.csv", stringsAsFactors = F)

head(input)


```


### Refromat the tibble
using stringr, dplyr, magrittr package
```{r}

library(magrittr)
library(stringr)
library(dplyr)

# pipe operators were used to simplify the transformation steps
# str_extract functoin was used to select strings based on regular expressions
# select function was used to pick the required columns for new dataframe

reformatted_df <- input %>%
  mutate(street = str_extract(address, "(\\b([[:alnum:]]+\\s+[[:alnum:]]+\\.*)[^,]+)"),
                 zip = str_extract(address, "[[:digit:]]+$")
               ) %>%
  select("brewery_name", "type", "website", "street", "state", "zip", "state_breweries")

  head(reformatted_df)

```


### Performing some analysis
using dplyr, magrittr package
```{r}

library(magrittr)
library(dplyr)

# counting brewery types
# group_by and tally functoins were used to count the brewery types
# arrange function was used to get clear idea of popular brewry type

count_by_type <- reformatted_df %>%
  group_by(type) %>%
  tally(name = "typecount") %>%
  arrange(.,desc(typecount))

count_by_type

```

### ggplot functionality added by Stephen Haslett as per Task 2 of the assignment - "Extend an Existing Example".
```{r, message=FALSE, warning=FALSE}
library(ggplot2)
brewery_type_plot <- ggplot(count_by_type, aes(x = reorder(type, - typecount), y = typecount)) +
    geom_bar(stat = "identity", fill = '#f28e1c') +
    labs(title = "Brewery Count by Type",
         x = "Brewery Type",
         y = "Count") +
    theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
    theme(panel.background = element_rect(fill = '#FFFFFF'))

brewery_type_plot
```
