---
title: "Tidyverse Part 1"
author: "C. Rosemond"
date: "November 2, 2019"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Library

```{r library, message=FALSE}
library(tidyverse)
```

## Data Set(s)
I selected two fivethirtyeight data sets: one that contains current Soccer Power Index (SPI) ratings and rankings for men's club teams and a second that contains match-by-match SPI ratings and forecasts back to 2016.

URL: https://github.com/fivethirtyeight/data/tree/master/soccer-spi

## readr - read_csv()
The readr package facilitates the reading in of 'rectangular' data like .csv files or other delimited files. Here, I use the read_csv() function to read in two data sets: the global rankings, or 'rankings', and the matches, or 'matches'.

```{r read_csv1, eval=TRUE, message=FALSE}
rankings <- read_csv('https://projects.fivethirtyeight.com/soccer-api/club/spi_global_rankings.csv')
matches <- read_csv('https://projects.fivethirtyeight.com/soccer-api/club/spi_matches.csv')
head(rankings)
tail(matches)
```


## dplyr - mutate()
The dplyr package provides a grammar for the manipulation of data--notably, in data frames or tibbles. Here, I use the mutate function to add a new column--a match ID--to the matches tibble.

```{r mutate, eval=TRUE}
matches <- mutate(matches, match = row_number())
matches <- matches[,c(23,1:22)]
head(matches)
```


## dplyr - select()
The select function from dplyr enables the selection of data frame columns by name or helper function. Here, I select and keep the first six columns ('match' through 'team2') from the matches tibble.

```{r select, eval=TRUE}
matches <- select(matches, match:team2)
head(matches)
```


## dplyr - filter()
The filter function from dplyr enables the subsetting of rows based on specified logical criteria. Here, I select matches that occurred from November 1st through November 7th.

```{r filter, eval=TRUE}
matches <- filter(matches, date >= '2019-11-01' & date <= '2019-11-07')
head(matches)
```


## tidyr - gather()
The tidyr package is designed to facilitate reshaping data. Here, I use the gather() function to reshape the matches tibble from wide to long format, gathering the separate team columns.

```{r gather, eval=TRUE}
matches <- matches %>% gather(-match, -date, -league_id, -league, key=team_number, value=name) %>% select(-team_number)
head(matches)
```


## dplyr - arrange()
The arrange function from dplyr enables the sorting of data based upon column values. Here, I arrange the matches tibble by match number.

```{r arrange, eval=TRUE}
matches <- arrange(matches, match)
head(matches)
```


## dplyr - left_join()
The left_join function works similarly to its SQL counterparts. I finish by using 'name' to merge the matches tibble with the rankings tibble, which contains club rankings and ratings as of November 7th.

```{r left_join, eval=TRUE}
merged <- dplyr::left_join(matches, rankings, by='name')
merged <- select(merged, -league.y, - off, -def)
head(merged)
```
## dplyr - rename()
We can use the rename function to clean up the 'league.x' column name
```{r}
merged <- rename(merged, 'league_name' = 'league.x')
head(merged)
```

## forcats - as_factor()

We can use forcats to take a look at the "league_name" field.  We first need to convert to a factor.

```{r}
merged$league_name = as_factor(merged$league_name)
table(merged$league_name)
```
## forcats - fct_infreq()

This table has no rhyme or reason.  We can use fct_infreq() to sort descending by frequency.

```{r}
league_table = as.data.frame(table(fct_infreq(merged$league_name)))
league_table = rename(league_table, 'league_name' = 'Var1')
league_table
```
## ggplot2 - geom_bar()

Now that we have our leagues in a nice order, we can use ggplot2 to plot the frequencies

```{r}
ggplot(league_table, aes(x=league_name, y = Freq)) + 
    geom_bar(stat="identity")


```

## ggplot2 - theme()

This is kind of a mess.  Let's use theme() to rotate the labels so that we can read them.

```{r}
ggplot(league_table, aes(x=league_name, y = Freq)) + 
    geom_bar(stat="identity") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))

```
