---
title: "Tidy Verse"
author: "Aaron Zalki"
date: "12/1/2019"
output: html_document
---

# {.tabset}

## Raw Drug-Use Data
### Background on the Data

I am tidying a fivethirtyeight dataset on drug use [Github](https://github.com/fivethirtyeight/data/tree/master/drug-use-by-age). I am transforming the columns of drugs (3 to 28) into rows, in order to perform analysis on the different drugs used by different age groups.
It is important to note that any column header in the dataset with the word 'use' is referring to a percentage of those in an age group who have used the specific drug in the past 12 months. 
Any column header in the dataset with the word 'frequency' is referring to the median number of times a user in an age group used the specific drug in the past 12 months.
For this analysis, I am only interested in the percentage the drug is used among different age groups.


### Loading Data

```{r raw data, include=TRUE}
library(RCurl)
raw_drug <- read.csv(text=getURL("https://raw.githubusercontent.com/fivethirtyeight/data/master/drug-use-by-age/drug-use-by-age.csv"), header=TRUE, stringsAsFactors = FALSE)
head (raw_drug)

```

## Tidy & Transform


I am using the packages tidyr and dplyr to change the raw data to its final form.
```{r transforming, include=TRUE}
library(tidyr)
library(dplyr)
drug_gather <- gather(raw_drug, "Drug", "Percent", 3:28)
drug_gather$Percent <- as.numeric(drug_gather$Percent)
drug_separate <- separate(drug_gather, "Drug", c("Drug", "Usage"))
drug_filter <- filter(drug_separate, Usage=='use')

names (drug_filter) <- c("Age", "Size", "Drug", "Usage", "Percent")

head(drug_filter)
```
## Analysis

### Teenagers

I am using the package ggplot to illustrate the differences found in drug use between teenagers and young adults.

```{r teen, include=TRUE}
library(ggplot2)
teen <- ggplot(subset(drug_filter, Age %in% c("13","14","15", "16", "17", "18","19")), 
aes(x=Drug, y=Percent, fill=Age)) + geom_bar(stat="identity", position=position_dodge()) + theme(axis.text.x = element_text(angle = 90, hjust = 1))

teen

```

### Young Adults

```{r young adult, include=TRUE}
young_adults <- ggplot(subset(drug_filter, Age %in% c("18","19","20","21","22-23","24-25")), 
aes(x=Drug, y=Percent, fill=Age)) + geom_bar(stat="identity", position=position_dodge()) + theme(axis.text.x = element_text(angle = 90, hjust = 1))


young_adults
```

