---
title: "DATA607_TidyVerse_assignment1"
author: "Lin Li"
date: "9/1/2019"
output: html_document
---

Create an Example.  Using one or more TidyVerse packages, and any dataset from fivethirtyeight.com or Kaggle, create a programming sample “vignette” that demonstrates how to use one or more of the capabilities of the selected TidyVerse package with your selected dataset. 

1. Get dataset from 538.

```{r}
weather <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/us-weather-history/KCLT.csv")
weather

```

2. tidyr separate function:
```{r}
library(tidyr)
weather2 <- weather %>% separate(date, c("year", "month", "day"), sep = "-")
weather2
```

3. dplyr select function:

```{r}
library(dplyr)
select(weather2, year, actual_mean_temp, record_min_temp, record_max_temp, record_precipitation)
```

4. dplyr filter (subsetting dataset)

```{r}
library(dplyr)
filter(weather2, year == "2014")
```

5. ggplot histogram for actual_mean_temp

```{r}
ggplot(weather2, aes(actual_mean_temp)) + geom_bar()
```

