---
title: "TIDYVERSE Part 2"
author: "Sadia Perveen"
date: "12/8/2019"
output: html_document
---
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library('knitr')
```

## Data Source

https://www.kaggle.com/ronitf/heart-disease-uci

## Load Data

```{r Load Data}
heart <- read_csv("https://raw.githubusercontent.com/forhadakbar/data607fall2019/master/Week%2014/heart.csv")
kable (head(heart))
```

## Capability 1.

`slice` capability tutorial

Description: Using slice we can select rows by specifying the row number.  
Usage: slice(data, …)  
Example: To select rows 6 to 12


```{r}
slice(heart, 6:12)
```

OR

To select rows 10 to 15, 18 and 299 to 302

```{r}
heart %>% slice(c(10:15, 18, 299:302))
```

## Capability 2.

`mutate` capability tutorial

Description: Creates new columns based on existing ones  
Usage: mutate(.data, …)  
Example: Let’s look at a ratio of resting blood pressure to cholesterol under a new column name “Ratio”

```{r}
#Ratio = trestbps/chol
heart %>% select(trestbps, chol) %>% mutate(Ratio = trestbps/chol)
```

## Capability 3.

`summarise` capability tutorial

Description: Applies functions that return results of length 1. Can perform multiple calculations in the same call.  
Usage: summarise(data, …)  
Example: Lets see mean and median cholesterol along with mean and median maximum heart rate achieved.

```{r}
heart %>% summarise(Avg_cholesteral = mean(chol), Median_cholesteral = median(chol), Avg_heartrate = mean(thalach), Median_heartrate = median(thalach))
```

## Capability 4.

`group_by` capability tutorial with tally

Description: Using group_by and tally we can find count of category member  
Usage: group_by(data, …)  
Example: Find the count by sex

```{r}
heart %>% group_by(sex) %>% tally()
```
## Capability 5.

`rename` capability

Description: Using rename function without changing the content of the data frame  
Rename column sex to M_F
Note this only affects the new dataframe - heart_new
```{r}
heart_new <- dplyr::rename(heart, M_F = sex)
names(heart_new)
```

Adding to Project 


1) Arraging the data by oldpeak
```{r}
arg_oldpeak <- dplyr::arrange(heart, oldpeak)
arg_oldpeak
```
Here we are comparing oldpeak and age 
```{r}
ggplot(heart, aes(x=oldpeak, y=age)) + geom_point()
```

