---
title: "DATA607_TidyVerse_assignment1"
author: "Lin Li"
date: "9/1/2019"
output: html_document
---

Create an Example.  Using one or more TidyVerse packages, and any dataset from fivethirtyeight.com or Kaggle, create a programming sample “vignette” that demonstrates how to use one or more of the capabilities of the selected TidyVerse package with your selected dataset. 

1. Get dataset from 538.

```{r}
weather <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/us-weather-history/KCLT.csv")
weather

```

2. tidyr separate function:
```{r}
library(tidyr)
weather2 <- weather %>% separate(date, c("year", "month", "day"), sep = "-")
weather2
```

3. dplyr select function:

```{r}
library(dplyr)
select(weather2, year, actual_mean_temp, record_min_temp, record_max_temp, record_precipitation)
```

4. dplyr filter (subsetting dataset)

```{r}
library(dplyr)
filter(weather2, year == "2014")
```

5. ggplot histogram for actual_mean_temp

```{r}
ggplot(weather2, aes(actual_mean_temp)) + geom_bar()
```

```{r}
weather3 <- weather2 %>%
 group_by(month) %>%
 summarise(avg = round(mean(actual_mean_temp), 1)) %>%
 arrange(as.numeric(month))
weather3$month <- as.character(weather3$month)
weather3$month[weather3$month == "1"] <- "January"
weather3$month[weather3$month == "2"] <- "February"
weather3$month[weather3$month == "3"] <- "March"
weather3$month[weather3$month == "4"] <- "April"
weather3$month[weather3$month == "5"] <- "May"
weather3$month[weather3$month == "6"] <- "June"
weather3$month[weather3$month == "7"] <- "July"
weather3$month[weather3$month == "8"] <- "August"
weather3$month[weather3$month == "9"] <- "September"
weather3$month[weather3$month == "10"] <- "October"
weather3$month[weather3$month == "11"] <- "November"
weather3$month[weather3$month == "12"] <- "December"
```

```{r}
library(ggplot2)
a <- ggplot(weather3, aes(month,avg))
a + geom_bar(stat = "identity") + scale_x_discrete(limits=weather3$month)
```
