---
title: "DATA607_TidyVerse_assignment1"
author: "Lin Li"
date: "9/1/2019"
output: html_document
---

Create an Example.  Using one or more TidyVerse packages, and any dataset from fivethirtyeight.com or Kaggle, create a programming sample “vignette” that demonstrates how to use one or more of the capabilities of the selected TidyVerse package with your selected dataset. 

1. Get dataset from 538.


```{r}
weather <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/us-weather-history/KCLT.csv")
weather
```

2. tidyr separate function:
```{r}
library(tidyr)
weather2 <- weather %>% separate(date, c("year", "month", "day"), sep = "-")
weather2
```

3. dplyr select function:

```{r}
library(dplyr)
select(weather2, year, actual_mean_temp, record_min_temp, record_max_temp, record_precipitation)
```

4. dplyr filter (subsetting dataset)

```{r}
library(dplyr)
filter(weather2, year == "2014")
```

5. ggplot histogram for actual_mean_temp

```{r}
library(ggplot2)
ggplot(weather2, aes(actual_mean_temp)) + geom_bar()
```


 
---
title: "DATA607_TidyVerse_assignment Part 2"
author: "Euclid Zhang"
date: "11/17/2019"
---
  
Update with new examples of the Tidyverse packages used above (tidyr, dplyr and ggplot2)
  
tidyr::gather - a function that transform the values from multiple columns into one column, with the original column names as keys stored in a new created column.  
For example, we can combine the actual_mean_temp, actual_min_temp and actual_max_temp into one columns:
```{r}
weather2 %>%
  select(year:actual_max_temp) %>% 
  gather("temp_type", "temperature", actual_mean_temp:actual_max_temp) %>% 
  head()
```
dplyr::group_by - group the rows by the values in column(s)  
dplyr::summarise - for each group determined by the group_by function, generate a summary value that is defined by a summary function such as mean()  
dplyr::arrange - arrange the order of the rows by the values in column(s)   
For example, we can group the rows by months and summarize the average of the mean temperature in each month:
```{r}
weather2 %>% 
  group_by(year, month) %>% 
  summarise(average = mean(actual_mean_temp)) %>% 
  arrange(as.integer(year),as.integer(month))
```

ggplot2 - the package provides various functions to plot multuple components on the same graph.  
For example, the geom_line function plots and connects each (x,y) pairs from provided input lists x and y.  
The geom_smooth function plots a smooth line accross the points.  
Labels can be created by the labs function.  
In the following example we plot the actual mean temperature of 365 days
```{r message=FALSE}
temp <- ggplot(weather2, aes(x=c(1:365), y=weather2$actual_mean_temp)) + 
  geom_line(color = "orange") +
  geom_smooth() +
  labs(title="Daily Mean Temperature from 07/01/2014 to 06/30/2015",
       x="Day",
       y="temperature (Fahrenheit)")
plot(temp)
```

We can also plot multiple lines on the same graph. For example, we can add the actual min temperature and max temperature to the same plot.
```{r message=FALSE}
temp <- ggplot(weather2, aes(x=c(1:365), y=weather2$actual_mean_temp)) + 
  geom_line(aes(y=weather2$actual_mean_temp),color = "orange") +
  geom_smooth(color = "black") +
  labs(title="Daily Temperature from 07/01/2014 to 06/30/2015",
       subtitle = "Orange = Mean, Red = Max, Blue = Min",
       x="Day",
       y="temperature (Fahrenheit)") +
  geom_line(aes(y=weather2$actual_min_temp),color = "blue") +
  geom_line(aes(y=weather2$actual_max_temp),color = "red")
plot(temp)
```

