---
title: "TidyVerse Assignment"
author: "Samuel I Kigamba"
date: "December 1, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

TidyVerse assignment, due 12/01, accepted until 12/08
In this assignment, you’ll practice collaborating around a code project with GitHub.  You could consider our collective work as building out a book of examples on how to use TidyVerse functions.

GitHub repository:  https://github.com/acatlin/FALL2019TIDYVERSE 

FiveThirtyEight.com datasets.

Kaggle datasets. 

You have two tasks:

Create an Example.  Using one or more TidyVerse packages, and any dataset from fivethirtyeight.com or Kaggle, create a programming sample “vignette” that demonstrates how to use one or more of the capabilities of the selected TidyVerse package with your selected dataset. (25 points)
Extend an Existing Example.  Using one of your classmate’s examples (as created above), extend his or her example with additional annotated code. (15 points)
You should clone the provided repository.  Once you have code to submit, you should make a pull request on the shared repository.  Minimally, you should be submitted .Rmd files; ideally, you should also submit an .md file and update the README.md file with your example.

After you’ve completed both parts of the assignment, please submit your GitHub handle name in the submission link provided in the week 1 folder!  This will let your instructor know that your work is ready to be graded.

You should complete both parts of the assignment and make your submission no later than the end of day on Sunday, December 1st.



```{r, message=FALSE, warning=FALSE}
library(stringr)
library(tidyr)
library(dplyr)
library(tidyverse)
```


# Data Source.

FiveThirtyEight.com datasets (https://data.fivethirtyeight.com/)

#Load Data directly from FiveThirtyEight.com

```{r, message=FALSE, warning=FALSE}

polls <- read_csv("https://projects.fivethirtyeight.com/polls-page/president_primary_polls.csv")

head(polls) #Print out first 5 rows to confirm that the data have been loaded correctly.

```



#Rename third and fourth columns using the rename() function.
```{r}
polls <- dplyr::rename(polls, cycle_year = cycle)
polls <- dplyr::rename(polls, state_name = state)

names(polls) #Column names after renaming the third and fourth columns.
```


Use the slice() function to select a few rows for example row 10 to row 15.
```{r}
polls_sliced <- polls %>% slice(c(10:15))
head(polls_sliced)
```

Use the groupby() function to group by category member and tally
```{r}
polls_tally <- polls %>% group_by(answer) %>% tally()
polls_tally

```

use the function glimpse() to view all the columns of the resulting data frame.
```{r}
dplyr::glimpse(polls_tally) 
```
```