---
title: "TidyVerseAssignment_Part1"
author: "Lin Li"
date: "12/1/2019"
output: html_document
---

## Task 1:
Create a programming sample “vignette” that demonstrates how to use one or more of the capabilities of the selected TidyVerse package with your selected dataset.

# My chosen dataset:
From fivethirtyeight: "Some people are too superstitious to have a baby on Friday the 13th"

Loading and exploring datasets:

```{r}
birth_NCHS <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/births/US_births_1994-2003_CDC_NCHS.csv")
birth_SSA <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/births/US_births_2000-2014_SSA.csv")
head(birth_NCHS)
```


# I would like to compare mean births on Fridays to other days to see if there is a difference in the dataset for 1994-2003.

```{r}
library(dplyr)

# getting mean births for each day of the week for both datasets:

NCHS_mean <- 
    birth_NCHS %>%
        group_by(day_of_week) %>% 
        summarise(mean_birthNCHS = mean(births))
NCHS_mean

SSA_mean <-
    birth_SSA %>% 
        group_by(day_of_week) %>% 
        summarise(mean_birthSSA = mean(births))
SSA_mean   
```

```{r}
library(ggplot2)
chart_birth_NCHS <- ggplot(data = NCHS_mean, aes(x=day_of_week, y=mean_birth)) +
    geom_bar(stat = "identity", width = 0.6, fill = "black")
chart_birth_NCHS

chart_birth_CDC <- ggplot(data = CDC_mean, aes(x=day_of_week, y= mean_birthCDC)) +
    geom_bar(stat = "identity", width = 0.6, fill = "red")
chart_birth_CDC
```


# Compare mean birth on Fridays between the two datasets, which represent two separate decades, 1994-2003 and 2000-2014.
```{r}
# join the two datasets:
library(dplyr)
joint_births <- left_join(NCHS_mean, SSA_mean, by = "day_of_week")
joint_births

```

The two datasets have similar births throughout the week. Surprisingly, Sunday has the lowest mean births for both datasets.