---
title: "Extending Avengers"
author: "Aaron Zalki"
date: "12/1/2019"
output: html_document
---


```{r}
library(tidyverse)
```

# I chose to work with the Avengers dataset 

```{r}
# Get dataset from URL from FiveThirtyEight github
Avengers <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv")
```

# I chose to demonstrate the dplyr package

```{r}
# Use filter to extract an individual avenger
Captain_America <- filter(Avengers, Name.Alias == "Steven Rogers")
Iron_Man <- filter(Avengers, Appearances == "3068")
Thor <- filter(Avengers, Name.Alias == "Thor Odinson")
Black_Widow <- filter(Avengers, Name.Alias == "Natalia Alianovna Romanova")
The_Hulk <- filter(Avengers, Name.Alias == "Robert Bruce Banner")
Hawkeye <- filter(Avengers, Name.Alias == "Clinton Francis Barton")
```

```{r}
# Use full_join to combine everything into one data frame
Avengers_2012 <- list(Captain_America, Iron_Man, Thor, Black_Widow, The_Hulk, Hawkeye) %>%
  reduce(full_join, by.x = "URL", by.y = "Appearances")
Avengers_2012
```

For my Tidyverse example I used the Avengers dataset from FiveThirtyEight. The dataset contains a list of members of the Avengers group from the Marvel comics. The list shows appearances of when they were first introduced in the comics and if/when they died and/or returned. I used the filter function from the dplyr package to extract every avenger from the 2012 Avengers film. I then put them into a list and used the full_join function from the dplyr function to put every avenger that I got from using the filter function into one data frame.


# Adding a plot to show the most apperances for characters with at least 2000 - Aaron Zalki

```{r adding a plot}
library (ggplot2)

appearances <- ggplot(subset(Avengers, Appearances > 2000), 
aes(x=Name.Alias, y=Appearances, fill=Appearances)) + geom_bar(stat="identity", position=position_dodge()) + theme(axis.text.x = element_text(angle = 90, hjust = 1))

appearances 
```

© 2019 GitHub, Inc.