---
title: "Tidyverse Assighment"
author: "Sadia Perveen"
date: "12/8/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this assignment, you’ll practice collaborating around a code project with GitHub.  You could consider our collective work as building out a book of examples on how to use TidyVerse functions.

GitHub repository:  https://github.com/acatlin/FALL2019TIDYVERSE 

FiveThirtyEight.com datasets.

Kaggle datasets. 

You have two tasks:

Create an Example.  Using one or more TidyVerse packages, and any dataset from fivethirtyeight.com or Kaggle, create a programming sample “vignette” that demonstrates how to use one or more of the capabilities of the selected TidyVerse package with your selected dataset. (25 points)
Extend an Existing Example.  Using one of your classmate’s examples (as created above), extend his or her example with additional annotated code. (15 points)
You should clone the provided repository.  Once you have code to submit, you should make a pull request on the shared repository.  Minimally, you should be submitted .Rmd files; ideally, you should also submit an .md file and update the README.md file with your example.

After you’ve completed both parts of the assignment, please submit your GitHub handle name in the submission link provided in the week 1 folder!  This will let your instructor know that your work is ready to be graded.

```{r}
install.packages("tidyverse")
library(tidyverse)
```



Import Data from FiveThirtyEight.com datasets.
```{r}
nutrition <- read.csv("https://raw.githubusercontent.com/sadia-perveen/TIDYVERSE/master/raw_anonymized_data.csv")
```

First we will use the dplyr package 

1)Here are are creating a new dataset where the 'MILKONCEREALFREQ' is arranged. 
```{r}
arg_milk <- dplyr::arrange(nutrition, MILKONCEREALFREQ)
arg_milk
```

2) Here are are creating a new dataset where the 'Cancer' is answered yes. 

```{r}
cancer_yes <- dplyr::filter(nutrition, cancer == "Yes")
cancer_yes
```

3) Selecting columns related to EGG consumption 

```{r}
dplyr::select(nutrition, starts_with("EGG"))
```

Now we will look at the ggplot2 package 

1) Here we are comparing YOGURTFREQ with EGGSFREQ 
```{r}
ggplot(nutrition, aes(x=EGGSFREQ, y=YOGURTFREQ)) + geom_point()
```

2) Barplot for cancer and GRITSFREQ
```{r}
 ggplot(nutrition, aes(x=cancer, y=GRITSFREQ)) +
 geom_col() 
```

