---
title: "TidyVerse_Assignment-Women World Cup 2019 Prediction"
author: "Don Padmaperuma (Geeth)"
date: "10/13/2019"
output: 
  html_document:
    toc: True
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(plotly)
```

## Assignment  

For this assignment I am using Women World Cup 2019 Prediction data set that I retrieved from [Fivethirtyeight](https://data.fivethirtyeight.com/) site to create a programming sample "vignette" that demonstrate how to use few of the main capabilities of the selected tidyverse package. 

### Introduction

This dataset shows the chance that each team will win, lose or tie every one of its matches. One important thing that this dataset include is the **SPI rating**, Soccer Power Index, how we turn those ratings into a forecast and how we calculate our in-game win probabilities.  
To generate our SPI ratings, we run through every past match in our database of women’s international matches — back to 1971 — evaluating the performance of both teams with four metrics:

1. The number of goals they scored.
2. The number of goals they scored, adjusted to account for red cards and the time and score of the match when each goal was scored.
3. The number of goals they were expected to score given the shots they took.
4. The number of goals they were expected to score given the nonshooting actions they took near the opposing team’s goal.


### Read .csv file  

```{r}
women_worldcup <- read_csv("https://projects.fivethirtyeight.com/soccer-api/international/2019/wwc_matches.csv")
```
  
```{r}
head(women_worldcup)
```

### Data transformation with dplyr  

In this section I will be using slice(), filter(), summerise(), count() functions that can be found under dplyr.  

#### Manipulate Cases

**Using slice option we can select the row numbers we want to display**  

```{r}
slice(women_worldcup, 10:20)
```
**filter function to extract rows that meet logical criteria**  

```{r}
filter(women_worldcup, women_worldcup$score1 > "3")
```

### Summarize Cases

**apply summarize function to columns to create new table of summary statistics**  

```{r}
women_worldcup %>% summarise(Avg_proj_score1 = mean(proj_score1), Median_proj_score1 = median(proj_score1), Avg_score1 = mean(score1), Median_score1 = median(score1))
```

**count number of rows each group defined by the variable**  

```{r}
#Team 1
count(women_worldcup, women_worldcup$team1, sort = TRUE)
```

```{r}
#Team 2
count(women_worldcup, women_worldcup$team2, sort = TRUE)
```

### Data Wrangling with tidyr

Under tidyr, There are four fundamental functions of data tyding.     
1. gather() takes multiple columns, and gathers them into key-value pairs: it makes “wide” data longer
2. spread() takes two columns (key & value) and spreads in to multiple columns, it makes “long” data wider
3. separate() splits a single column into multiple columns
4. unite() combines multiple columns into a single column

#### Reshaping data

Creating a new data set with selected columns. Then gather some columns.   

```{r}
#Eliminating few columns from the original data set and creating a new data set. 
wwc_19 <- select(women_worldcup, -c(date, league_id, league, xg1, xg2, nsxg1, nsxg2, prob1, prob2, probtie))
```

```{r}
df <- wwc_19 %>%
  gather(Teams, team_names, team1:team2)
df
```

I am going to create a new two columns by combining scores and spi rank.  This new data set will help me for my data visualization part as well.  

```{r}
only_scores <- wwc_19 %>%
  unite(score_comb, score1, score2, sep = "-") %>%
  unite(spi_comb, spi1, spi2, sep = "-")
only_scores
```

### Data Visualization - heatmaps  

I am using ggplot function to show a heatmap like plot. 

#### Visualization of scores    

```{r}
#Graph to show the scores  

g <- ggplot(wwc_19, aes(x = team1, y = team2, fill = score1-score2)) +
  geom_label(aes(label = paste(score1, score2, sep = "-")), fill = "white") +
  scale_x_discrete(limits = levels(wwc_19$team1), position = "bottom") +
  scale_y_discrete(limits = rev(levels(wwc_19$team2))) +
  theme (axis.text.x = element_text(angle = 60, hjust = 1))

g
```  

```{r}
only_scores <- wwc_19 %>%
  unite(score_comb, score1, score2, sep = "-") %>%
  unite(spi_comb, spi1, spi2, sep = "-")

heatmap <- ggplot(only_scores, mapping = aes(x = team2, y = team1, fill = score_comb))+
  geom_tile()+
  theme (axis.text.x = element_text(angle = 60, hjust = 1))
ggplotly(heatmap)
```

#### Visualization of SPI (Soccer Power Index) Rating  


```{r}
heatmap <- ggplot(only_scores, mapping = aes(x = team2, y = team1, fill = spi_comb))+
  geom_tile()+
  theme (axis.text.x = element_text(angle = 60, hjust = 1))
ggplotly(heatmap)
```
