---
title: "Tidyverse"
author: "Donny Lofland"
date: "November 23, 2019"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Library & Helpers

```{r library, message=FALSE}
library(tidyverse)
library(lubridate)

# This helper function will load URL data and store it locally 
# in a data/ folder for future loading without having to 
# redownload the file every time.  The function will create the
# data/ folder if it doesn't already exist.
# 
# Note that we are using the readr write_csv() method
load_cached_url <- function(filename, url) {
  dir.create(file.path('.', 'data'), showWarnings = FALSE)
  fqfn <- paste('./data', filename, sep = '/')

  # Download the file and save it, if we don't have it yet.  
  if(!file.exists(fqfn)) {
    data <- read_csv(url)
    write_csv(data, fqfn)
  }

  # Guess column specification.
  fqfn_spec <- spec_csv(fqfn)
  
  # Identify date columns.
  # Assumption: Column name ends with "date".
  date_cols <- names(fqfn_spec$cols)[str_detect(names(fqfn_spec$cols),
                                     pattern = "date$")]
  
  # Remove one of the date column names, in order to illustrate later the
  # alternative of specifying datatype after reading, too.
  date_cols <- date_cols[-1]

  # Set specification explicitly for date columns.
  for (column in date_cols) {
    fqfn_spec$cols[[column]] <- col_date("%m/%d/%Y")  
  }
  
  # Read file, setting datatypes for all columns.
  data <- read_csv(fqfn, col_types = fqfn_spec)
  
  return(as_tibble(data))
}
```

## Data Set(s)
FiveThirtyEight has aggregated polling data from a number of sources tracking Trump's approval rating since January 2017.  In this programming vignette, we will use TidyVerse packages to load, manipulate, summarize and plot Trump's approval rating over time. 

URL: https://github.com/fivethirtyeight/data/tree/master/trump-approval-ratings

There are two CSV files of interest:
- [Approved Polls](https://projects.fivethirtyeight.com/trump-approval-data/approval_polllist.csv)
- [Poll Results](https://projects.fivethirtyeight.com/trump-approval-data/approval_topline.csv)

## readr - read_csv() and spec_csv() methods
The readr package provides methods for loading and saving data from both the local file system and from url's where data is hosted online.  We will use the read_csv() and write_csv() method to load our CSV data and save it to a local cache.

spec_csv() allows us to specify datatypes for the read. If identification of datatypes can be generalized, we can program the coercion of datatype at run time of the read. In this illustration, the names of date columns for each input file all end with "date". Therefore, we can specify the date datatype upon reading the files.

Otherwise, it is necessary to coerce datatypes after reading. In order to illustrate this later, we code this vignette intentionally to miss one of the date columns.

```{r read_csv1, eval=TRUE, message=FALSE}

polls <- load_cached_url(
  'trump-approval-ratings-polls.csv', 
  'https://projects.fivethirtyeight.com/trump-approval-data/approval_polllist.csv'
  )


results <- load_cached_url(
  'trump-approval-ratings-results.csv',
  'https://projects.fivethirtyeight.com/trump-approval-data/approval_topline.csv')

head(polls)
head(results)
```

## lubridate

In order to demonstrate lubridate, our file reading function intentionally omitted setting the datatype of the first date columns. Let's now make sure all date columns are treated as a date format using lubridate's mdy(). 

```{r}
date_cols <- c('modeldate', 'startdate', 'enddate', 'createddate')

polls$modeldate   <- mdy(polls$modeldate)
results$modeldate   <- mdy(results$modeldate)
```

## dplyr - filter() & select()

The dplyr:filter() function allows us to filter rows much like a 'where' statement in SQL.  Here, we will only consider polls of Voters.  The dplyr:select() function will limit which columns are returned.  Here, we only want the date, approval and disapproval values.  The dplyr:arrange() functions allows us to sort the rows by model date.  We leverage the dyplr:gather() function to convert our tibble from wide to long format which aids with plotting.

```{r filter, eval=TRUE}
# example filtering, selecting an sorting data
res <- results %>%
  filter(subgroup == 'Voters') %>%
  select(modeldate, approve_estimate, disapprove_estimate) %>%
  arrange(modeldate)

# example creating a new column
res <- res %>%
  mutate(gap = (approve_estimate - disapprove_estimate))

# example converting to long format for plotting
res <- res %>% 
  gather('type', 'value', -modeldate) 

head(res)
```

### tidyr 1.0
The most recent version of `tidyr` uses a new set of verbs. Specifically,
instead of `gather` and `spread`, there is `pivot_longer` and `pivot_wider`.
It is more sensitive than the older version, and won't allow pivoting different
variable types to longer. The older version used to cast everything to the most
general type, which in this case would be character. So we well compare
distributing the `president` and `subgroup` variable over `modeldate`.

```{r pivot}
res1 <- res %>% select(president, subgroup, modeldate) %>% gather('type', 'value', -modeldate)
res2 <- %>% select(president, subgroup, modeldate) %>% pivot_longer(cols = -modeldate, names_to = "type", values_to = "value")
all.equal(res1, res2)
```

## ggplot2

Let's see how Trump's approval and disapproval ratings have varied over time.  We also plot the gap, calculated as approval minus disapproval.  We see that very early into his term, Trump's disapproval rating exceeded his approval rating and he has maintained a negative gap throughout his presidency.  

```{r}
# Multiple line plot
ggplot(res, aes(x = modeldate, y = value)) + 
  geom_area(aes(color = type, fill = type), 
            alpha = 0.2, position = position_dodge(0.8)) +
  scale_color_manual(values = c("#00FF00", "#ff0000", "#ff0000")) +
  scale_fill_manual(values = c("#00FF00", "#ff0000", "#E7B800")) 
```

